---
title: "Darwin Core Mapping"
subtitle: "FOURCAST legacy data publication"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
library(ggplot2)        # for map
library(mapview)
```

# Read source data

Create a data frame `input_data1` & `input_data2` from the source data:



```{r}
input_data1 <- read_excel(here::here("datasets", "Collembola", "data", "raw", "collembola_BF_plots_mei1997_orig.xlsx"))

```

```{r}
input_data2 <- read_excel(here::here("datasets", "Collembola", "data", "raw", "EB_Site_Table.xlsx"))

```


Pivot original dataframe `pivot longer`


```{r}

input_data1 %<>% 
pivot_longer(cols = !scientificName, names_to = "Site", values_to = "count")
```

Join both dataframes 

```{r}
input_data <- inner_join(input_data1, input_data2, by = NULL)
```


Preview data:

```{r}
input_data %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")
```

```{r}
input_data %>% head(n = 5)
```



## Lambert to Decimals

### Keep original x,Y as verbatimcoordinates



### Create a spacial dataframe

Create spatial dataframe and define the CRS (31370 is Belgian Lambert 72)
We indicate the x,y columns as coordinates and define the coordinate system


## Transform the data in WGS84

Now we transform the Lambert 72 in WGS84


### Create a dataframe with column decimalLatitude & decimalLongitude

In order to add the decimal coordinates into the original df we create an x_y dataframe


# Create extra dataframe to work with 



# Remove geometry from dataframe 'input_interim_df'



### Bind columns x_y and input_interim_df

Join both df's to have the decimalLat & Lon


### Integrate x-y coordinates in original dataframe

We can safely substitute input_interim 


# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_data
```

## remove obsolete columns


# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.org/10.15468/XXXXX")
```

### add row numbers to genrate occurrenceID unique


```{r}
occurrence %<>% mutate(ID_row = row_number())
```


### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "INBO") # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
#occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "INBO") # e.g. "INBO"
```

### CollectionCode


```{r}
occurrence %<>% mutate(dwc_collectionCode = "FOURCAST")
```



### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Forest soil classification - Collembola, Flanders, Belgium") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

### ownerInstitutionCode

```{r}
occurrence %<>% mutate(dwc_ownerInstitutionCode = "INBO | Ugent | UAntwerp")
```

### CollectionCode


```{r}
occurrence %<>% mutate(dwc_collectionCode = "FOURCAST")
```


The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### informationWithHeld

### dataGeneralizations

### occurrenceID

```{r}
occurrence %<>% mutate(occurrenceID = row_number()) %<>%
                mutate(dwc_occurrenceID = str_c("SOIL:COLLEMBOLA:", occurrenceID))
  
```



### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = 'Luc De Bruyn | Wim Jacobs | Frans Janssens')
```

### individualCount

```{r}
occurrence %<>%  mutate(dwc_individualCount = count) %<>%
                 mutate(dwc_individualCount = ifelse(is.na(dwc_individualCount), 0, dwc_individualCount))
                
```


```{r}


#occurrence$individualCount %<>% replace_na(0)
```

### organismQuantity




### organismQuentityType



### sex

### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(dwc_occurrenceStatus = case_when(dwc_individualCount > 0 ~ "Present",
                                        dwc_individualCount == 0 ~ "Absent",
                                        is.na(dwc_individualCount) ~ "Absent" ))

```


### lifeStage

### behavior

### occurrenceRemarks

##event

### eventDate


```{r}
occurrence %<>% mutate(dwc_eventDate = '1997-05')

```

#event
## eventID

Need to create an eventID based on name and site ID

```{r}
occurrence %<>% mutate(dwc_eventID = str_c(Site, dwc_eventDate))
```

## habitat



## samplingProtocol


```{r}
occurrence %<>% mutate(dwc_samplingProtocol = 'pitfall traps') 

```

### samplingEffort

```{r}
occurrence %<>% mutate(dwc_samplingEffort = 'pitfall traps')

```



# Location

### locationID

Need to create an eventID based on name and site ID

```{r}
occurrence %<>% mutate(dwc_licationID = Site)
```

```{r}
occurrence %<>%
  rename(dwc_decimalLongitude = Longitude) %>%
  rename(dwc_decimalLatitude = Latitude) %>%
  mutate(dwc_geodeticDatum = "WGS84") %>%
  mutate(dwc_coordinateUncertaintyInMeters = "18")
  
```

```{r}
occurrence %<>%
  mutate(dwc_decimalLongitude = round(dwc_decimalLongitude, digits = 5)) %>%
  mutate(dwc_decimalLatitude = round(dwc_decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(dwc_decimalLatitude = as.character(format(dwc_decimalLatitude, nsmall = 5))) %>%
   mutate(dwc_decimalLongitude = as.character(format(dwc_decimalLongitude, nsmall = 5)))
```

### continent

```{r}
occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality

### verbatimLocality

```{r}
occurrence %<>% mutate(dwc_verbatimLocality = 'Site Name') 
```

### verbatimcoordinates

### verbatimLatitude

### verbatimLongitude

### verbatimcoordinatesystem

```{r}
occurrence %<>% mutate(verbatimcoordinatesystem = "Lambert coordinates") # 
```

### verbatimSRS

```{r}
occurrence %<>% mutate(verbatimSRS = "Belgian Datum 1972")
```

### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

# geologicalContext

## formation


```{r}
occurrence %<>% mutate(dwc_formation = `Observation Region`)

```


### eventID

Need to create an eventID based on name and site ID

```{r}
occurrence %<>% mutate(dwc_formation = str_c(Site, dwc_eventDate))
```
# identification

## identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = 'Luc De Bruyn | Wim Jacobs | Frans Janssens')
```

#taxon

### scientificName 

```{r}
occurrence %<>% mutate(dwc_scientificName = scientificName)
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### scientificNameAuthorship


### taxonRank

```{r}
occurrence %<>% mutate(taxonRank = "species")
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(nomenclaturalCode = "ICZN") # e.g. "ICZN"
```



## Post-processing

Only keep the Darwin Core columns:

```{r}
occurrence %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

Preview data:

```{r}
occurrence %>% head()
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

Define the order of the output columns


Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "Collembola", "data", "processed", "occurrence.csv"), na = "")
```


Visualize data on a map

```{r}
occurrenceMap <- occurrence
```

```{r}
occurrenceMap$decimalLatitude <- as.numeric(occurrenceMap$decimalLatitude)
```

```{r}
occurrenceMap$decimalLongitude <- as.numeric(occurrenceMap$decimalLongitude)
```

```{r}
head(occurrenceMap)
```



```{r}


head(occurrence)

mapview(occurrenceMap, xcol = "decimalLongitude", ycol = "decimalLatitude", crs = 4269, grid = FALSE)

```

